---
title: "03-regression"
author: "nilseling"
date: "2020-05-12"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Simple linear regression

We're going to work with the cars dataset

```{r}
str(cars)

head(cars)
tail(cars)
```

### Simple visualisation


```{r viz}
library(tidyverse)

cars %>%
  ggplot() + geom_point(aes(x = speed, y = dist))
```


### Fitting the model

```{r lm-function}
stop_dist_model <- lm(dist ~ speed, data = cars)

stop_dist_model

summary(stop_dist_model)
```

### Intro to broom package

```{r broom}
library(broom)

broom::augment(stop_dist_model)
broom::tidy(stop_dist_model)
broom::glance(stop_dist_model)
```

We can now visulize the model fit:

```{r viz-2}
augment(stop_dist_model) %>%
  ggplot() + geom_point(aes(speed, dist)) +
  geom_line(aes(speed, .fitted), colour = "red", lwd = 2) +
  geom_segment(aes(speed, dist, xend = speed, yend = .fitted)) +
  geom_label(data = glance(stop_dist_model), aes(x = 5, y = 120, label = paste0("R2: ", round(r.squared, 2))))
```

**Calculate the regression coefficients**

```{r}
x <- cars$speed
y <- cars$dist

Sxy <- sum((x - mean(x)) * (y - mean(y)))
Sxx <- sum((x - mean(x)) ^ 2)

beta_hat_1 <- Sxy / Sxx
beta_hat_0 <- mean(y) - beta_hat_1 * mean(x)

c(beta_hat_0, beta_hat_1)
stop_dist_model
```

**Calculating the standard error**

```{r SE}
cur_resid <- residuals(stop_dist_model)
cur_resid

manual_resid <- y - (beta_hat_0 + beta_hat_1 * x)
manual_resid

all.equal(as.numeric(cur_resid), manual_resid)
```

Calculate model variance

```{r}
se_2 <- sum(manual_resid ^ 2) / (nrow(cars) - 2)
se_2

sqrt(se_2)

summary(stop_dist_model)
```

**Calculate a variance decomposition**

```{r}
y_hat <- beta_hat_0 + beta_hat_1 * x

SST <- sum((y - mean(y)) ^ 2)
SSE <- sum((y - y_hat) ^ 2)
SSReg <- sum((y_hat - mean(y)) ^ 2)

R2 <- SSReg / SST
R2

summary(stop_dist_model)
```

## Inference for SLR

**Calculating the standard errors of the regression coefficients**

```{r}
Sxx <- sum((x - mean(x)) ^ 2)
s_e <- sqrt(se_2)
n <- nrow(cars)

SE_beta_hat_0 <- s_e * sqrt(1 / n + mean(x) ^ 2 / Sxx)
SE_beta_hat_1 <- s_e /  sqrt(Sxx)

c(SE_beta_hat_0, SE_beta_hat_1)

tidy(stop_dist_model)
```

**Calculate the t statistic**

```{r t-statistic}
t_beta_hat_0 <- (beta_hat_0 - 0) / SE_beta_hat_0
t_beta_hat_1 <- (beta_hat_1 - 0) / SE_beta_hat_1

c(t_beta_hat_0, t_beta_hat_1)
tidy(stop_dist_model)
```

**Calculate the p-value**

```{r p-value}
p_beta_hat_0 <- 2 * pt(abs(t_beta_hat_0), df = nrow(cars) - 2, lower.tail = FALSE)
p_beta_hat_1 <- 2 * pt(abs(t_beta_hat_1), df = nrow(cars) - 2, lower.tail = FALSE)
```









