---
title: "04-plotting"
author: "nilseling"
date: "2020-05-19"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

In todays session, we will go through a number of common plot types and will learn how to generate them using `ggplot2` and other packages of the "ggverse".
For simplicity, I will again use the `iris` dataset and will use this to highlight a couple of different plots.

```{r}
data(iris)
library(ggplot2)
```

ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPING>),
    stat = <STAT>, position = <POSITION>) + 
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION> +
  <THEME_FUNCTION>
  
## Scatterplot

Create scatterplot `Sepal.Width` versus `Sepal.Length`:

```{r scatter-1}
p1 <- ggplot(data = iris) +
  geom_point(mapping = aes(x = Sepal.Length, y = Sepal.Width, colour = Species))
p1
```

Quickplot `qplot`

```{r qplot}
qplot(data = iris, x = Sepal.Width, y = Sepal.Length, colour = Species, geom = "point")
```

### Add another layer

Add a linear regression line per group

```{r geom_smooth}
p1 <- p1 + geom_smooth(mapping = aes(x = Sepal.Length, y = Sepal.Width, colour = Species), 
                       method = "lm", formula = "y ~ x")
p1
```

### Facetting

```{r facetting}
p1 <- p1 + facet_wrap(. ~ Species)
p1
```

Use the NPG colour scheme

```{r change-colours}
library(ggsci)
p1 <- p1 + scale_colour_npg()
p1
```

Create a colour vector:

```{r RColorBrewer}
library(RColorBrewer)
display.brewer.all(type = "qual", colorblindFriendly = TRUE)
cur_pal <- brewer.pal(n = 3, name = "Set2")
colour_vec <- c(setosa = cur_pal[1],
                versicolor = cur_pal[2],
                virginica = cur_pal[3])

p1 <- p1 + scale_color_manual(values = colour_vec)
p1
```

Start from scratch

```{r scatter-2}
p1 <- ggplot(data = iris) +
  geom_point(mapping = aes(x = Sepal.Length, y = Sepal.Width, fill = Species),
             shape = 21, size = 2) +
  geom_smooth(mapping = aes(x = Sepal.Length, y = Sepal.Width, colour = Species), 
                       method = "lm", formula = "y ~ x") +
  facet_wrap(. ~ Species, labeller = labeller(Species = c("setosa" = "Setosa",
                                                          "versicolor" = "Versicolor",
                                                          "virginica" = "Virginica"))) +
  scale_color_manual(values = colour_vec, name = "Iris species", labels = c("Setosa", "Versicolor", "Virginica")) +
  scale_fill_manual(values = colour_vec, name = "Iris species", labels = c("Setosa", "Versicolor", "Virginica")) +
  xlab("Sepal Length") + ylab("Sepal Width") + 
  theme(panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        legend.position = c(0.67, 0.8),
        legend.background = element_rect(colour = "black"),
        axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(size = 12),
        strip.background = element_blank(),
        strip.text = element_text(colour = "black", size = 12))
p1
```

## Plotting a PCA

We can use the custom PCA function.

```{r load-custumPCA}
source("code/helper.R")
cur_pca <- prcomp(iris[,1:4], scale. = TRUE)
p2 <- my_plotPCA(cur_pca, colour_by = iris$Species)
p2
```

Change the theme and other:

```{r change-appearance}
p2 <- p2 + scale_colour_manual(values = colour_vec, name = "Iris species",
                    labels = c("Setosa", "Versicolor", "Virginica")) +
  coord_fixed() + 
  theme(panel.background = element_blank(),
        legend.background = element_rect(colour = "black"),
        legend.position = c(1, 1), 
        legend.justification = c("right", "top"),
        axis.line = element_line(),
        axis.ticks = element_line(colour = "black"),
        axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 12))
p2
```

## Boxplot

I will plot the distribution of `Sepal.Length` across the different species:

```{r boxplot}
set.seed(12345)
p3 <- ggplot(data = iris) +
  geom_boxplot(mapping = aes(x = Species, y = Sepal.Length, fill = Species), 
               outlier.colour = NA, notch = TRUE) +
  geom_jitter(aes(x = Species, y = Sepal.Length), width = 0.1, alpha = 0.7) +
  scale_fill_manual(values = colour_vec, name = "Iris species",
                    labels = c("Setosa", "Versicolor", "Virginica")) +
  scale_x_discrete(breaks = c("setosa", "versicolor", "virginica"),
                   labels = c("Setosa", "Versicolor", "Virginica")) +
  xlab("") +
  ylab("Sepal Length") +
  theme(panel.background = element_blank(),
        legend.background = element_rect(colour = "black"),
        legend.position = c(0.05, 0.95), 
        legend.justification = c("left", "top"),
        axis.line = element_line(),
        axis.ticks = element_line(colour = "black"),
        axis.text.y = element_text(colour = "black", size = 9),
        axis.text.x = element_text(colour = "black", size = 12),
        axis.title = element_text(colour = "black", size = 12))
p3
```

## Barplot

Caclulate the mean, sd, se and ci of the 4 different features per species:

```{r}
library(tidyverse)
cur_dat <- iris %>% 
  pivot_longer(cols = colnames(iris)[1:4]) %>%
  group_by(Species, name) %>%
  summarise(mean_val = mean(value),
            sd_val = sd(value),
            n_val = length(value),
            se_val = sd_val / sqrt(n_val),
            ci_t = qt(0.975, df = n_val - 1) * se_val)

cur_dat
```

Display this in barplot:

```{r barplot}
cur_dat$Measurement <- factor(cur_dat$name, levels = c("Petal.Width", "Petal.Length", "Sepal.Width", "Sepal.Length"))
p4 <- ggplot(cur_dat, mapping = aes(x = Species, y = mean_val, fill = Measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean_val - ci_t, ymax = mean_val + ci_t), 
                position = position_dodge(width = 0.9), width = 0.2) +
  scale_fill_brewer(palette = "Paired",
                    labels = c("Petal Width", "Petal Length", "Sepal Width", "Sepal Length")) +
  labs(x = "", y = "Measured value [cm]") +
  scale_x_discrete(breaks = c("setosa", "versicolor", "virginica"),
                   labels = c("Setosa", "Versicolor", "Virginica")) +
    theme(panel.background = element_blank(),
        legend.background = element_rect(colour = "black"),
        legend.position = c(0.05, 0.95), 
        legend.justification = c("left", "top"),
        axis.line = element_line(),
        axis.ticks = element_line(colour = "black"),
        axis.text.y = element_text(colour = "black", size = 9),
        axis.text.x = element_text(colour = "black", size = 12, angle = 45, hjust = 1),
        axis.title = element_text(colour = "black", size = 12))
p4
```

## Heatmap

rows: all features, col: all samples

```{r}
library(pheatmap)

cur_dat <- t(iris[1:4])
colnames(cur_dat) <- rownames(iris)
p5 <- pheatmap(cur_dat, scale = "row", color = colorRampPalette(c("dark blue", "white", "dark red"))(100),
               annotation_col = data.frame(row.names = colnames(cur_dat),
                                           Species = iris$Species),
               annotation_colors = list(Species = colour_vec),
               show_colnames = FALSE,
               labels_row = sub("\\.", " ", rownames(cur_dat)))
```

## Stitching it together

```{r cowplot}
library(cowplot)

upper_row <- p1
middle_row <- plot_grid(p2 + theme(legend.title = element_text(size = 10),
                                   legend.text = element_text(size = 8),
                                   legend.key.size = unit(0.1, "cm"),
                                   legend.position = c(1, 1.3)), 
                        p3 + theme(legend.title = element_text(size = 10),
                                   legend.text = element_text(size = 8),
                                   legend.key.width = unit(0.3, "cm")), 
                        p4 + theme(legend.title = element_text(size = 8),
                                   legend.text = element_text(size = 6),
                                   legend.position = c(0.05, 1.1),
                                   legend.key.size = unit(0.3, "cm")), 
                        ncol = 3, labels = c("B", "C", "D"))
lower_row <- p5$gtable

final_figure <- plot_grid(upper_row, middle_row, lower_row, labels = c("A", "", "E"), nrow = 3)

ggsave(filename = "docs/final_figures/Fig_1.pdf", plot = final_figure, width = 210, height = 290, units = "mm")
```







